---
title: "employee characteristics"
author: "Martin Monkman"
date: "04/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages
library(tidyverse)
library(lubridate)

library(glue)
library(here)
```

## Employee characteristics


The CSV file "df_HR_main.csv" contains the following variables:

* emp_id -- a unique identification number for each employee

* start_date -- the date the employee started their employment with CR25

* four variables based on the designated groups specified in Canada's [_Employment Equity Act_](https://laws-lois.justice.gc.ca/eng/acts/E-5.401/), where "TRUE" indicates the the employee is a member of that group

  - women
  
  - Aboriginal peoples
  
  - persons with disabilities
  
  - members of visible minorities (meaning "persons, other than Aboriginal peoples, who are non-Caucasian in race or non-white in colour").



## employee frame

establish frame with employee ID

```{r}

df_HR <- as_tibble(nums <- sprintf('%0.3d', 1:999)) 
  
df_HR <-
df_HR %>% mutate(emp_id = as.character(glue("ID{value}"))) %>% 
  select(emp_id)


df_HR

```

start date

function from https://stackoverflow.com/questions/21502332/generating-random-dates/21502397#21502397

```{r date_function}

rdate <- function(n = 1,
                  min = paste0(format(Sys.Date(), '%Y'), '-01-01'),
                  max = lubridate::today(),
                  sort = TRUE) {

  dates <- sample(seq(as.Date(min), as.Date(max), by = "day"), n, replace = TRUE)
  if (sort == TRUE) {
    sort(dates)
  } else {
    dates
  }

}

```


```{r}
set.seed(42)

#rdate(n = 10, max = "2020-11-22")

df_HR <- 
  df_HR %>% 
  mutate(
#    date_of_birth = rdate(n = n(), min = "1960-01-01", max = "2000-01-01", sort = FALSE),
    start_date = rdate(n = n(), min = "2010-01-01", max = "2020-12-31")
    )

df_HR


# write base version of HR frame with employee ID and start date
write_rds(df_HR, here("data", "df_HR.rds"))

```

## assign equity group

women: 50% 
visible minority: 25%
aboriginal: 10%
disability: 15%

```{r}

set.seed(42)

df_HR_main <- df_HR %>% 
  mutate(equity_woman = sample(c(TRUE, FALSE), size = n(), prob = c(.5, .5), replace = TRUE),
         equity_visible_minority = sample(c(TRUE, FALSE), size = n(), prob = c(.25, .75), replace = TRUE),
         equity_aboriginal = sample(c(TRUE, FALSE), size = n(), prob = c(.10, .90), replace = TRUE),
         equity_disability = sample(c(TRUE, FALSE), size = n(), prob = c(.15, .85), replace = TRUE)) 

df_HR_main

df_HR_main %>% 
  group_by(equity_visible_minority, equity_aboriginal) %>% 
  tally()

write_csv(df_HR_main, here("data", "df_HR_main.csv"))

```


