---
title: "hr staff turnover"
output: html_document
---

Scripts to generate a fake HR relational database


## The scenario: 

The fictional wholesale company, CR25 (for "Classic Rock 25"), has 999 employees. The data files cover the period from 2010-01-01, when the company first opened its doors, through to 2020-12-31. 


### employee characteristics

The CSV file "df_HR_main.csv" contains the following variables:

* emp_id -- a unique identification number for each employee

* start_date -- the date the employee started their employment with CR25

* four variables based on the designated groups specified in Canada's [_Employment Equity Act_](https://laws-lois.justice.gc.ca/eng/acts/E-5.401/), where "TRUE" indicates the the employee is a member of that group

  - women
  
  - Aboriginal peoples
  
  - persons with disabilities
  
  - members of visible minorities (meaning "persons, other than Aboriginal peoples, who are non-Caucasian in race or non-white in colour").


### transaction history

Some of the employees in this company have been with the company since it started on 2010-01-01, and others started recently.

There are 5 occupations in the company, labelled A (for Apprentice) to E (for Executive), and each has more responsibility (and therefore pay) than the one below. Employees can start at any one of the occupations, and are eligible for a promotion after 90 days on the job (but some people take longer than that). 

Each occupation has 3 steps (numbered 1 to 3); everyone starts at step 1, and moves to step 2 after 1 year in that occupation, and step 3 after 2 years. 

The second CSV file "df_HR_transaction.csv" contains the promotion history for each employee. The variables are:

* emp_id -- a unique identification number for each employee

* transaction_num -- the hiring and promotion event categories

* transaction_date -- the date of the hiring and promotion event

* occupation -- the five occupation categories

* occ_step -- the annual increment steps


(The code for the creation of these data files is here:  https://github.com/MonkmanMH/hr-example-data)




```{r setup}

library(tidyverse)
library(glue)
library(lubridate)
library(magrittr)
```


## employee frame

establish frame with employee ID

```{r}

df_HR <- as_tibble(nums <- sprintf('%0.3d', 1:999)) 
  
df_HR <-
df_HR %>% mutate(emp_id = as.character(glue("ID{value}"))) %>% 
  select(emp_id)


df_HR

```

start date

function from https://stackoverflow.com/questions/21502332/generating-random-dates/21502397#21502397

```{r date_function}

rdate <- function(n = 1,
                  min = paste0(format(Sys.Date(), '%Y'), '-01-01'),
                  max = lubridate::today(),
                  sort = TRUE) {

  dates <- sample(seq(as.Date(min), as.Date(max), by = "day"), n, replace = TRUE)
  if (sort == TRUE) {
    sort(dates)
  } else {
    dates
  }

}

```


```{r}
set.seed(42)

#rdate(n = 10, max = "2020-11-22")

df_HR <- 
  df_HR %>% 
  mutate(
#    date_of_birth = rdate(n = n(), min = "1960-01-01", max = "2000-01-01", sort = FALSE),
    start_date = rdate(n = n(), min = "2010-01-01")
    )

df_HR

```

equity group

```{r}

set.seed(42)

df_HR_main <- df_HR %>% 
  mutate(equity_woman = sample(c(TRUE, FALSE), size = n(), prob = c(.5, .5), replace = TRUE),
         equity_visible_minority = sample(c(TRUE, FALSE), size = n(), prob = c(.25, .75), replace = TRUE),
         equity_aboriginal = sample(c(TRUE, FALSE), size = n(), prob = c(.10, .90), replace = TRUE),
         equity_disability = sample(c(TRUE, FALSE), size = n(), prob = c(.15, .85), replace = TRUE)) 

df_HR_main

df_HR_main %>% 
  group_by(equity_visible_minority, equity_aboriginal) %>% 
  tally()

write_csv(df_HR_main, "df_HR_main.csv")

```


---

## transaction table

To base table, add three rows with transaction dates then pivot

Note: some individuals will have only one transaction (start date), others will have two, and some will have three


```{r}

# test runs to work out how to add random amount to existing date without going past today

var_date = ymd("2019-01-01")

today() - var_date

# use today - variable date to get number of days, then 
# multiply by a random value from the uniform distribution (runif) to get 
# a fractional number of days to add
var_date + ((today() - var_date) * runif(1))

```


```{r}
# test runs
df_HR %>% 
  mutate(trans1 = (start_date + ((today() - start_date) * runif(n())))) %>% 
  mutate(trans_time = trans1 - start_date,
         service = today() - start_date,
         as.integer(trans_time) / as.integer(service))

df_HR %>% 
  mutate(trans1 = (start_date + ((today() - start_date) * runif(n())))) %>% 
  mutate(trans2 = (trans1 + ((today() - trans1) * runif(n())))) %>% 
  mutate(trans3 = (trans2 + ((today() - trans2) * runif(n())))) %>% 
    mutate(trans_time = trans3 - trans2,
         service = today() - start_date,
         as.integer(trans_time) / as.integer(service))

df_HR %>% 
  mutate(trans1 = (start_date + ((today() - start_date) * runif(n())))) %>% 
  mutate(trans2 = (trans1 + ((today() - trans1) * runif(n())))) %>% 
  mutate(trans3 = (trans2 + ((today() - trans2) * runif(n())))) %>% 
    mutate(trans_time = trans3 - trans2,
         service = today() - start_date,
         as.integer(trans_time) / as.integer(service))

```

finals

```{r}
set.seed(42)

df_HR_transaction_dates <- 
  df_HR %>% 
  mutate(trans1 = (start_date + ((today() - start_date) * runif(n())))) %>% 
  mutate(trans2 = (trans1 + ((today() - trans1) * runif(n())))) 
#%>% 
#  mutate(trans3 = (trans2 + ((today() - trans2) * runif(n()))))

df_HR_transaction_dates

# remove small gaps by converting to NA

df_HR_transaction_dates <-
df_HR_transaction_dates %>% 
  # trans1 
  mutate(
    trans1 = case_when(
             as.integer(trans1 - start_date) < 90 ~ NA_Date_,
             TRUE ~ trans1)
  ) %>% 
  # trans2 
  mutate(
    trans2 = case_when(
             is.na(trans1) ~ NA_Date_,
             as.integer(trans2 - trans1) < 90 ~ NA_Date_,
             TRUE ~ trans2)
  ) 
#%>% 
  # trans3
#  mutate(
#    trans3 = case_when(
#             is.na(trans2) ~ NA_Date_,
#             as.integer(trans3 - trans2) < 90 ~ NA_Date_,
#             TRUE ~ trans3)
#  ) 

df_HR_transaction_dates  

```

pivot longer

```{r}

df_HR_transaction_dates %<>% 
  pivot_longer(!emp_id, names_to = "transaction_num", values_to = "transaction_date")


df_HR_transaction_dates

```




### occupational group


occupation -- start with transaction 2 -- if !na, then sample from 3 - 4 - 5 (so they could have entered as 1 - 2 - 3 and got 2 promotions)

then work backwards

```{r}
df_HR_transaction_dates <- df_HR_transaction_dates %>% 
  mutate(
    occ_grp_num = case_when(
      # fill NA
      transaction_num == "trans2" & is.na(transaction_date) ~ NA_integer_,
      transaction_num == "trans1" & is.na(transaction_date) ~ NA_integer_,
      # fill trans2
      transaction_num == "trans2" ~ sample(c(3L, 4L, 5L), 
                                           size = n(), replace = TRUE),
      TRUE ~ 991L)
  )

df_HR_transaction_dates <- df_HR_transaction_dates %>% 
  mutate(
    # fill trans1
    occ_grp_num = case_when(
      is.na(occ_grp_num) ~ NA_integer_,
      occ_grp_num <= 5 ~ occ_grp_num,
      transaction_num == "trans1" & is.na(lead(occ_grp_num, 1)) ~ sample(c(2L, 3L, 4L, 5L), 
                                                                    size = n(), replace = TRUE),
      transaction_num == "trans1" ~ lead(occ_grp_num, 1) - 1L,
      TRUE ~ 991L
  )
  )

df_HR_transaction_dates <- 
  df_HR_transaction_dates %>% 
  mutate(
    # fill start
    occ_grp_num = case_when(
      is.na(occ_grp_num) ~ NA_integer_,
      occ_grp_num <= 5 ~ occ_grp_num,
      transaction_num == "start_date" & is.na(lead(occ_grp_num, 1)) ~ sample(c(1L, 2L, 3L, 4L, 5L), 
                                                                    size = n(), replace = TRUE),
      transaction_num == "start_date" ~ lead(occ_grp_num, 1) - 1L,
      TRUE ~ 999L
  )
  )

  

```

recode to alphabetical

```{r}
df_HR_transaction_dates <-
  df_HR_transaction_dates %>% 
  mutate(
    occupation = case_when(
      is.na(occ_grp_num) ~ NA_character_,
      occ_grp_num == 1 ~ "A",
      occ_grp_num == 2 ~ "B",
      occ_grp_num == 3 ~ "C",
      occ_grp_num == 4 ~ "D",
      occ_grp_num == 5 ~ "E",
      TRUE ~ "999"
  )
  )

df_HR_transaction_dates

```  

quick tally of total

```{r}

df_HR_transaction_dates %>% 
  group_by(transaction_num, occupation) %>% 
  tally()

```



for steps, today - transaction 2 > 2 * 365 = 3 (i.e. two years past, gone to step 3), 


```{r}

df_HR_transaction_dates <-
df_HR_transaction_dates %>% 
  mutate(
    yr_in_occ = case_when(
      is.na(transaction_date) ~ NA_real_,
      transaction_num == "trans2" ~ 
        as.numeric(interval(transaction_date, today()) %/% years(1)) + 1,
      
      (transaction_num == "trans1" | transaction_num == "start_date") 
      & is.na(lead(transaction_date, 1)) ~ 
        as.numeric(interval(transaction_date, today()) %/% years(1)) + 1,

      transaction_num == "trans1" | transaction_num == "start_date" ~
        as.numeric(interval(transaction_date, lead(transaction_date, 1)) %/% years(1)) + 1,

            TRUE ~ 999
  ))

df_HR_transaction_dates

tail(df_HR_transaction_dates)

```


some quick tallys

```{r}

df_HR_transaction_dates %>% 
  group_by(occupation) %>% 
  summarise(avg_years = mean(yr_in_occ))

```

# recode to steps

```{r}
df_HR_transaction_dates <-
  df_HR_transaction_dates %>% 
  mutate(
    occ_step = case_when(
      is.na(yr_in_occ) ~ NA_integer_,
      yr_in_occ >= 3 ~ 3L,
      yr_in_occ == 2 ~ 2L,
      yr_in_occ == 1 ~ 1L,
      TRUE ~ NA_integer_
  )
  )

df_HR_transaction_dates

```  


select columns

```{r}
df_HR_transaction <-
df_HR_transaction_dates %>% 
  select(-c(occ_grp_num, yr_in_occ))

df_HR_transaction

```




write file

```{r}
  
write_csv(df_HR_transaction, "df_HR_transaction.csv")

```



```{r}
# rounding date differences

(today() - ymd("2019-11-22"))

interval(ymd("2018-11-22"), today()) %/% years(1)


```

